{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-11 thesis/thesis.tex
revise the concl per daves wishes

Concurrency testing is, in a way, a microcosm for research in general:
an exponentially large problem
that invariably thwarts they whoever try to solve it ``perfectly'';
it bears fruit only to one who can find their own human-oriented meaning in partial results and compromises.
I speak of course literally on one side of this analogy, figuratively on the other.
This thesis has demonstrated such meaning for stateless model checking,
choosing perhaps one of the most challenging demographics of programmers in need of concurrency testing,
the undergraduate operating systems student,
as its target audience.
Along the way, it justified its techniques' applicability to problems beyond CMU's walls as well
by showing positive results at other universities
and by addressing new concurrency models developed by industry in just the last decade.
% also by comparing the preemption bounds of quicksand eval's bugs to prior work in ICB
% and finding them to actually be skewed to DEEPER bugs than in the "real" "world", so take that
As I write this I am also collaborating with 15-410 staff
to integrate Landslide as an official part of the course grading infrastructure,
so in some sense the na\"{i}ve wish to make an ``automatic TA'' that drew my attention to grad school to begin with
has also been fulfilled.

For as long as people have written science fiction books,
we have fantasized about robots that can make any complex intuitive decision a human can
in addition to the superhuman arithmetic ability that comes standard with silicon.
Recent trends in machine learning have pursued this aesthetic,
using statistics and pattern recognition to provide convincing output even on never-before-seen inputs,
but few such systems are designed to also alert the user when their output confidence is low.
% appropriate for serving ads (penalty for being wrong -- no click)
% appropriate for drug synthesis (clinical trial vets the output)
% not appropriate for... govt surveillance o.o;
% not appropriate for program verification
As tempting as it is to fantasize an oracle capable of telling for certain whether your program is free of concurrency bugs,
that would be a disservice to both user and machine:
the human skill that artificial intelligence should really try to emulate is
knowing its limits and asking for help when it gets confused.
% ways humans cooperate to make landslide better:
% - test case design incl. preemption sealing
% - abstraction reduction
% - manual inspection of data races

It's become fashionable among my peers lately,
whenever a new exploit is announced in the news,
to declare that security is doomed,
that computers were a mistake,
that we should all retire and become llama farmers,
and so on.
There is no doubt that we are burdened with massive technical debt from the mistakes made
during computer science's infancy.
Nevertheless,
safety-oriented programming languages and type systems are growing ever more popular,
and meanwhile formal verification techniques to check existing code,
written in the old unsafe ways, grow ever more powerful.
Landslide's ability to reach impressionable student minds is my little contribution in this long battle.
I have hope that we may one day live in peace with our silicon creations,
understanding and respecting their limitations just as they complement ours.
This thesis has been a profoundly meaningful journey for me and I hope you enjoyed reading it.

>>>>

Concurrency testing is, in a way, a microcosm for research in general:
an exponentially large problem
that invariably thwarts they whoever try to solve it ``perfectly'';
it bears fruit only to one who can find their own human-oriented meaning in partial results and compromises.
I speak of course literally on one side of this analogy, figuratively on the other.
This thesis has demonstrated such meaning for stateless model checking,
choosing perhaps one of the most challenging demographics of programmers in need of concurrency testing,
the undergraduate operating systems student,
as its target audience,
\revision{ultimately helping 73\% of participating students to find bugs in their code,
56\% to fix any such bugs,
and 34\% to fix all such bugs,
before their project deadline.}
Along the way, I justified \revisionminor{Landslide's} applicability to problems beyond CMU's walls as well
by showing positive results at other universities,
and by addressing a new concurrency model developed by \revisionminor{the hardware} industry in just the last decade.
% also by comparing the preemption bounds of quicksand eval's bugs to prior work in ICB
% and finding them to actually be skewed to DEEPER bugs than in the "real" "world", so take that
As I write this I am also collaborating with 15-410 staff
to integrate Landslide as an official part of the course grading infrastructure,
so in some sense the na\"{i}ve wish to make an ``automatic TA'' that drew my attention to grad school to begin with
has also been fulfilled.

\revision{Let us revisit the first half of the thesis statement,
which in \cref{sec:thestatement} I mentioned would serve as the work's overarching theme:
that by
{\em combining theoretically-founded automatic reduction techniques and user-informed heuristic ones,}
stateless model checking can meet human testing needs.
How have each of the major contributions of this work embodied this theme?
%Let me also opine briefly on which of this work's contributions constitute lasting lessons
%that any future research building hereupon should heed.
Let me also briefly address:
What lasting lessons have we learned from each,
which any future research building hereupon would be wise to heed?

\begin{itemize}
	\item In \Cref{chap:quicksand},
		the soundness of data-race-driven Iterative Deepening
		combined with Quicksand's reliance on state space estimation
		to effectively balance formal verification with heuristic bug-finding speed
		better than any prior work.
		% 未来は、
		Even as modern programming languages such as Rust \cite{rust-book}
		threaten to obsolete the data-race preemption point proof
		by statically forbidding data races to begin with,
		Iterative Deepening's general approach remains relevant alongside ICB \cite{chess-icb},
		each providing a different tradeoff between bug-finding and completion time,
		and future model checkers should offer a mixture of different kinds of partial verification results
		depending on the user's wishes.
	\item In \Cref{chap:education},
		the behind-the-scenes yield-loop blocking algorithm
		(itself admittedly heuristic)
		allowed Landslide to automatically instrument all student implementations
		with very few false positives,
		and exposing partial results such as data race reports
		was shown to sometimes help students reach deeper understanding.
		% 未来は、
		Although high-level undergraduate operating systems courses represent
		a very small fraction of all the world's programmers,
		let alone all those in higher education,
		Landslide's adventures in 15-410 argue that computer science classes
		should more intentionally teach principled debugging techniques and tools
		rather than let students struggle to reach correctness independently. %on their own.
		Also, any future debugging tool should heed my students' survey feedback
		and take friendly user interface design more seriously than I did with Landslide.
	\item In \Cref{chap:tm},
		the equivalence of HTM's interface to failure injection and global locks
		allowed Landslide to avoid both performance overhead and state space explosion when simulating aborts,
		and the heuristic use of abort code reduction and abstraction reduction
		allowed verifying algorithms and data structures up to thread/iteration counts previously infeasible.
		% retry set reduction, obv; it also applies to failure injection as well
		% even when synch is complicated, defining abstraction boundaries and verifying your shit separately
		% is ,,good
		% - even if you have to be careful that you're "proving the right thing"
		% 未来は、
		Emulating complex execution semantics with simpler, already-understood primitives,
		and the retry set reduction,
		both allowing substantial yet sound state space reduction,
		are the obvious take-aways for any future model checker wishing to tackle HTM.
		I expect the latter should easily apply to other opportunities for failure-injection beyond transactions as well,
		such as memory allocation, filesystems, and distributed systems.
		More in general,
		the verification section demonstrated the value of compartmentalizing one's verification efforts,
		checking modules separately against a common, well-understood interface
		to reach yet further up the exponential cliff.
\end{itemize}
}

For as long as people have written science fiction books,
we have fantasized about robots that can make any complex intuitive decision a human can
in addition to the superhuman arithmetic ability that comes standard with silicon.
Recent trends in machine learning have pursued this aesthetic,
using statistics and pattern recognition to provide convincing output even on never-before-seen inputs,
but few such systems are designed to also alert the user when their output confidence is low.
% appropriate for serving ads (penalty for being wrong -- no click)
% appropriate for drug synthesis (clinical trial vets the output)
% not appropriate for... govt surveillance o.o;
% not appropriate for program verification
As tempting as it is to fantasize an oracle capable of telling for certain whether your program is free of concurrency bugs,
that would be a disservice to both user and machine:
the human skill that artificial intelligence should really try to emulate is
knowing its limits and asking for help when it gets confused.
% ways humans cooperate to make landslide better:
% - test case design incl. preemption sealing
% - abstraction reduction
% - manual inspection of data races
\revision{Here I have demonstrated that stateless model checking can realistically fulfill this ideal,
and I envision a future of concurrency verification where % when
human and program may cooperate to debug and verify even more complex software than ever before.}

It's become fashionable among my peers lately,
whenever a new exploit is announced in the news,
to declare that security is doomed,
that computers were a mistake,
that we should all retire and become llama farmers,
and so on.
There is no doubt that we are burdened with massive technical debt from the mistakes made
during computer science's infancy.
Nevertheless,
safety-oriented programming languages and type systems are growing ever more popular,
and meanwhile formal verification techniques to check existing code,
written in the old unsafe ways, grow ever more powerful.
Landslide's ability to reach impressionable student minds is my little contribution in this long battle.
I have hope that we may one day live in peace with our silicon creations,
understanding and respecting their limitations just as they complement ours.
%This thesis has been a profoundly meaningful journey for me and I hope you enjoyed reading it.
